---
editor_options:
  chunk_output_type: console
---

NOTES:

-   This file is *not* included in the [source package](https://r-pkgs.org/package-structure-state.html#source-package) because of the [`.nopurl` suffix in its
    filename](https://rpkg.dev/pkgpurl/reference/purl_rmd.html#-rmd-files-excluded-from-purling).

-   The chunks below have to be manually executed in order to regenerate the package data.

# Define data

NOTES:

-   Although the datasets below are saved as ["internal data" in `R/sysdata.rda`](https://r-pkgs.org/data.html#data-sysdata), some of them might still be
    exported and documented (in the main `.Rmd` document), something [not explicitly
    mentioned](https://coolbutuseless.github.io/2018/12/10/r-packages-internal-and-external-data/) in the book [R
    Packages](https://r-pkgs.org/data.html#data-data).

## `item_keys`

Supported qstnr specification part keys.

TODO:

-   Extract as much of the info as possible from the JSON Schemas!

```{r}
item_keys <-
  readr::read_csv(file = "data-raw/item_keys.csv",
                  na = "",
                  col_types = "cccllccc") %>%
  dplyr::mutate(components = stringr::str_split(string = components,
                                                pattern = ","),
                default_val = purrr::map2(.x = default,
                                          .y = type,
                                          .f = ~ {
                                            if (is.na(.x)) {
                                              eval(parse(text = glue::glue("{.y}()")))
                                            } else {
                                              eval(parse(text = .x))
                                            }
                                          }),
                has_default = purrr::map_lgl(default_val,
                                             length)) %>%
  dplyr::select(-default)
```

## `notice`

```{r}
notice <- list(mandatory = list(de = "Diese Frage muss zwingend beantwortet und kann nicht \u00fcbersprungen werden.",
                                en = "This question is mandatory and cannot be skipped.",
                                fr = "Il faut obligatoirement répondre à cette question et on ne peut pas la sauter.",
                                it = "A questa domanda è obbligatorio rispondere e non può essere saltata."),
               multiple_answers = list(de = "Bei dieser Frage sind mehrere Antworten m\u00f6glich.",
                                       en = "More than one answer is possible for this question.",
                                       fr = "Il y a plus d'une réponse possible à cette question.",
                                       it = "Per questa domanda è possibile più di una risposta."))
```

## `qstnr_cols`

Overall canonical order of all the cols the different `qstnr` representations can have.

```{r}
qstnr_cols <-
  item_keys$key |>
  setdiff("i") %>%
  c("order",
    "group",
    "lang",
    .)

qstnr_cols <- c(qstnr_cols[1:which(qstnr_cols == "values")],
                "values_int",
                "values_targets",
                "value_order",
                "value",
                "value_int",
                "value_targets",
                qstnr_cols[which(qstnr_cols == "value_scale"):length(qstnr_cols)])
```

## `spec_parts`

TODO:

-   Create `question_blocks.toml` with sensible sample data (a block "demographics" etc.).

-   Rethink whether we actually need/want to incl the default config in the package, or the pkgdown-based URL sourcing is enough.

```{r}
spec_parts <-
  fs::dir_ls(path = "data-raw/qstnr/configs",
             type = "directory") %>%
  rlang::set_names(nm = fs::path_file(.)) %>%
  purrr::map(~ fs::dir_ls(path = .x,
                          type = "file",
                          glob = "*.toml") %>%
               rlang::set_names(nm = fs::path_ext_remove(fs::path_file(.))) %>%
               purrr::map(pal::toml_read))

# TODO: remove this once items are moved to pkg fokus
spec_parts$default$items_TODO <- NULL
```

## `survey_config_keys`

```{r}
survey_config_keys <- names(jsonlite::fromJSON(txt = "data-raw/qstnr/schemas/qstnr-survey.schema.json")$properties$qstnr$properties)
```

## `value_scales*`

```{r}
value_scales <- jsonlite::fromJSON(txt = "data-raw/qstnr/schemas/subschemas/qstnr-item-value_scale.schema.json")$enum
value_scales_strict <- setdiff(value_scales,
                               "none")
```

# Write data

Save all the small data objects as a single internal file `R/sysdata.rda`. Note that when documenting them, they must be explicitly `@export`ed to be available
to package users.

```{r}
usethis::use_data(item_keys,
                  notice,
                  qstnr_cols,
                  spec_parts,
                  survey_config_keys,
                  value_scales,
                  value_scales_strict,
                  internal = TRUE,
                  overwrite = TRUE,
                  compress = "xz",
                  version = 3L)
```
